---
title: "DiscrepancyDiscourseAnalysis"
author: "Jean-Francois Chartier"
date: "12 février 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#imports
```{r}
library(magrittr)
library(Matrix)
library(quanteda)
library(proxy)
library(ggplot2)
```


#functions
```{r}
#function to project query in latent semantic space
  #'@param matrixV, a term-dim matrix m*k, of m terms previouslly modelleled with a SVD of k latent dimensions
  #'@param sigularValues, the k singular values of the train SVD
  #'@param newData, a new document-term matrix n*m to be projected in the latent space of k dimensions 
predictFromTrainSvdModel<-function(matrixV, singularValues, newData)
{
  call <- match.call()
  tsa <-  newData %*% matrixV %*% solve(diag((singularValues)))
  result <- list(docs_newspace = tsa)
  return (result)
}

#function to create a unit normed vector
normVector <- function(x) 
{
  if(sum(x)==0)
    return (x)
  else 
    return (x / sqrt(sum(x^2)))
  
}
#function to norm many vectors
normRowVectors<-function(m){
  t(apply(m, MARGIN = 1, FUN = function(x) normVector(x)))
}

#function to compute the dot product between 2 vectors
#note that when dot-product is applied to 2 unit vectors, it is the same as computing the cosine metric
dotProduct <- function(x, y) sum(x * y)
```


#Load data
```{r}
myMatrix = readRDS("2Banks_sparseMatrix-2019-02-25.rds")
    
myData = readRDS("dataForDashboardWithoutRareWords-2018-11-05.rds")
myData[,c(9,10,11)]=NULL    
myReducedMatrix=readRDS("2Banks_approxReducedMatrix-2019-02-12.rds")
latentNormedDocSpace = as.matrix(myReducedMatrix$u %*% solve(diag((myReducedMatrix$d)))) %>% normRowVectors()

selected.segment.by.threshold=readRDS("selected.segment.by.threshold.rds")


```

#1. Baseline analysis

Baseline is the discrepancy between all the text segments in the corpus, regardless of any concept.
Discrepancy between the toobigtofail concept need to be compared to this baseline in order to know if the observed difference between banks is specific to the studied concept or if it is only a difference between corpus  

##compute similarity between all segments
```{r}

years=seq(from=1999, to=2018, by=1)
avg.sim.by.year=data.frame(year=integer(length(years)), avg.sim=double(length(years)), std=double(length(years)), stringsAsFactors = F)

for(i in 1: length(years)){
  year_i=years[i]
  print(year_i)
  x=myData$bank=="BoE" & myData$year==year_i
  y=myData$bank=="ECB" & myData$year==year_i
  #x=myData%>%subset(., bank=="BoE")%>%.$year==year_i
  #y=myData%>%subset(., bank=="ECB")%>%.$year==year_i
  
  #equivalent to row-wise dot-product, but should be faster
  simil.betw.bank=tcrossprod(latentNormedDocSpace[x,], y = latentNormedDocSpace[y,])

  
  #simil.betw.bank=proxy::simil(x=latentNormedDocSpace[x,], y = latentNormedDocSpace[y,], by_rows=T, method=dotProduct, convert_distances = FALSE)
  
  avg.sim.by.year$year[i]=year_i
  avg.sim.by.year$avg.sim[i]=mean(simil.betw.bank)
  avg.sim.by.year$std[i]=sd(simil.betw.bank)
  
}

```

##plot averaged similarity distribution
```{r}
#plot(avg.sim.by.year$avg.sim, type = "l")
p=ggplot(avg.sim.by.year, aes(x=year, y=avg.sim))+
  geom_line()+
  geom_errorbar(aes(ymin=avg.sim-std, ymax=avg.sim+std), colour="black", width=.1)+
  geom_point(size=3)

p+labs(title="Evolution of discourse discrepancy between Central Banks\n measered with the cosine metric", x="Year", y = "Averaged cosine", subtitle="error bar = standard deviation ")
```

#2. toobigtofail concept analysis
Perform the analysis only on relevant segments where the studied concept is saillant 

##select min threshold
Select minimal cosine between the query and the segments

```{r}
minThreshold=0.18
j=which(colnames(selected.segment.by.threshold)==minThreshold)
relevantId=selected.segment.by.threshold[,j]
myQueryData=myData[relevantId, ]
```

##Frequency distribution by bank

Retrieve the proportion of segments by bank where the studied concept is at work 

```{r}
all.boe.freq.year=myData$year %>% subset(., myQueryData$bank=="BoE")%>%table(.)

all.ecb.freq.year=myData$year %>% subset(., myQueryData$bank=="ECB")%>%table(.)

boe.freq.year=myQueryData$year %>% subset(., myQueryData$bank=="BoE")%>%table(.)

ecb.freq.year=myQueryData$year %>% subset(., myQueryData$bank=="ECB")%>%table(.)

years=seq(from=1999, to=2018, by=1)
freq.by.year=data.frame(year=years, england=rep(0,length(years)), europe=rep(0,length(years)), stringsAsFactors = F)

for(i in 1: length(years)){
  #print(i)
  y=years[i]
  
  freq.by.year$england[i]=boe.freq.year[which(names(boe.freq.year)==y)] %>%ifelse(test = length(.)>0, yes = ., no = 0)%>%divide_by(., all.boe.freq.year[which(names(all.boe.freq.year)==y)])
  
  
  freq.by.year$europe[i]=ecb.freq.year[which(names(ecb.freq.year)==y)]%>%ifelse(test = length(.)>0, yes = ., no = 0)%>%divide_by(., all.ecb.freq.year[which(names(all.ecb.freq.year)==y)])
  
  
}

```

##plot frequency distribution by bank
```{r}
library(reshape2)
x=melt(freq.by.year, id.vars=1) %>%set_colnames(., c("year", "bank", "n.segment"))

ggplot(x, aes(x = year, y = n.segment, color = bank)) +
  theme_bw() + geom_line(size=2)+
  ylab("Proportion of segments")+xlab("Years")+
  #theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Proportion of semantically relevant segments by year and bank")
```


##compute similarity between relevant segments

Compute the averaged cosine between segments by year and by bank

```{r}
relevant.vectors=latentNormedDocSpace[relevantId,]

#concept.simil.betw.bank=proxy::simil(x=relevant.vectors[myQueryData$bank=="BoE",], y = relevant.vectors[myQueryData$bank=="ECB",], by_rows=T, method=dotProduct, convert_distances = FALSE)

#equivalent to row-wise dot-product, but should be faster
concept.simil.betw.bank=tcrossprod(x=relevant.vectors[myQueryData$bank=="BoE",], y = relevant.vectors[myQueryData$bank=="ECB",])

years=seq(from=1999, to=2018, by=1)
avg.concept.sim.year=data.frame(year=integer(length(years)), avg.sim=double(length(years)), std=double(length(years)), stringsAsFactors = F)

for(i in 1: length(years)){
  year_i=years[i]
  
  x=myQueryData%>%subset(., bank=="BoE")%>%.$year==year_i
  y=myQueryData%>%subset(., bank=="ECB")%>%.$year==year_i
  
  sub.simil.year=concept.simil.betw.bank[x, y]
  avg.concept.sim.year$year[i]=year_i
  avg.concept.sim.year$avg.sim[i]=mean(sub.simil.year)
  avg.concept.sim.year$std[i]=sd(sub.simil.year)
  
}

```

##plot averaged similarity distribution
```{r}
#plot(avg.sim.by.year$avg.sim, type = "l")
p=ggplot(avg.concept.sim.year, aes(x=year, y=avg.sim))+
  geom_line()+
  geom_errorbar(aes(ymin=avg.sim-std, ymax=avg.sim+std), colour="black", width=.1)+
  geom_point(size=3)

p+labs(title="Evolution of the toobigtoofail concept discrepancy between Central Banks\n as measured with the cosine metric", x="Year", y = "Averaged cosine", subtitle="error bar = standard deviation ")
```


#3. Analysis of specificities
get lexical specificities of relevant segments from the two banks
```{r}
#select bank
bank="ECB" #"BoE"

docvars(myMatrix, "ECB")<-((myData$bank=="ECB") & relevantId)
subDFM<-dfm_subset(myMatrix, relevantId)

docvars(subDFM, "BoE")=subDFM@docvars$ECB==F
    
#calculer les specificites
specificites = quanteda::textstat_keyness(x=subDFM, target=subDFM@docvars$ECB==T, measure="chi2", sort=TRUE)
    
#filter NA
specificites=specificites[is.na(specificites$chi2)==F,]
    
```

##Plot specificities
```{r}

p=0.05
specificites=specificites[specificites$p<p]

#set number of specificities to plot 
nSpec=40
#set min word count. Since we used the chi2, we should keep only words above 10
min_count=10
quanteda::textplot_keyness(x=specificites, n=nSpec, show_legend = T, color = c("darkblue", "darkred"), labelsize=2.5, min_count=min_count, labelcolor="black", margin = 0.2)+
  ggplot2::labs(title = "Lexical Specificities of Central Banks", x = "Chi-2", y = "Lexical specificities", color = "Central Banks")+theme(plot.margin = unit(c(1,4,1,2), "cm"))+ theme(legend.text = element_text(colour="black"))+
  scale_color_manual(labels = c("ECB", "BoE"), values = c("darkblue", "darkred"))

```



